---
title: "Untitled"
author: "John Tipton"
date: "6/6/2017"
output: html_document
---



```{r, warning=FALSE, message=FALSE}
library(BayesComposition)
library(ggplot2)
N <- 500
d <- 4
```



## Gaussian Simulation

```{r, warning=FALSE, message=FALSE}
dat <- sim_compositional_data(N=N, d=d, likelihood = "gaussian",
                              additive_correlation = TRUE)
simPlotData <- data.frame(species=as.factor(rep(1:d, each=N)),
                          count=c(dat$y),
                          depth=rep(dat$X, times=d), alpha=c(dat$alpha))

gsim1 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25)

gsim2 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth by species") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25) +
  facet_wrap( ~ species, ncol = 4)
multiplot(gsim1, gsim2, cols=1)
```



## multi-logit simulation using basis functions without overdispersion

```{r}
dat <- sim_compositional_data(N=N, d=d, likelihood = "multi-logit",
                              additive_correlation = FALSE,
                              multiplicative_correlation = TRUE)

y_dens <- matrix(0, N, d)
p_alpha <- matrix(0, N, d)
for (i in 1:N) {
  y_dens[i, ] <- dat$y[i, ] / sum(dat$y[i, ])
  p_alpha[i, ] <- exp(dat$mu + dat$zeta[i, ]) / sum(exp(dat$mu + dat$zeta[i, ]))
}

simPlotData <- data.frame(species=as.factor(rep(1:d, each=N)),
                          count=c(y_dens),
                          depth=rep(dat$X, times=d), alpha=c(p_alpha))

gsim1 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25)

gsim2 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth by species") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25) +
  facet_wrap( ~ species, ncol = 4)
multiplot(gsim1, gsim2, cols=1)
```



## multi-logit simulation using basis functions with overdispersion

```{r}
dat <- sim_compositional_data(N=N, d=d, likelihood = "multi-logit",
                              additive_correlation = TRUE,
                              multiplicative_correlation = TRUE)

y_dens <- matrix(0, N, d)
p_alpha <- matrix(0, N, d)
for (i in 1:N) {
  y_dens[i, ] <- dat$y[i, ] / sum(dat$y[i, ])
  p_alpha[i, ] <- exp(dat$mu + dat$zeta[i, ]) / sum(exp(dat$mu + dat$zeta[i, ]))
}

simPlotData <- data.frame(species=as.factor(rep(1:d, each=N)),
                          count=c(y_dens),
                          depth=rep(dat$X, times=d), alpha=c(p_alpha))

gsim1 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25)

gsim2 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth by species") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25) +
  facet_wrap( ~ species, ncol = 4)
multiplot(gsim1, gsim2, cols=1)
```



## multi-logit simulation using Gaussian process without overdispersion

```{r}
dat <- sim_compositional_data(N=N, d=d, likelihood = "multi-logit",
                              function_type = "gaussian-process",
                              correlation_function = "gaussian",
                              additive_correlation = FALSE,
                              multiplicative_correlation = TRUE)

y_dens <- matrix(0, N, d)
p_alpha <- matrix(0, N, d)
for (i in 1:N) {
  y_dens[i, ] <- dat$y[i, ] / sum(dat$y[i, ])
  p_alpha[i, ] <- exp(dat$mu + dat$zeta[i, ]) / sum(exp(dat$mu + dat$zeta[i, ]))
}

simPlotData <- data.frame(species=as.factor(rep(1:d, each=N)),
                          count=c(y_dens),
                          depth=rep(dat$X, times=d), alpha=c(p_alpha))

gsim1 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25)

gsim2 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth by species") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25) +
  facet_wrap( ~ species, ncol = 4)
multiplot(gsim1, gsim2, cols=1)
```



## multi-logit simulation using Gaussian process without overdispersion

```{r}
dat <- sim_compositional_data(N=N, d=d, likelihood = "multi-logit",
                              function_type = "gaussian-process",
                              correlation_function = "gaussian",
                              additive_correlation = TRUE,
                              multiplicative_correlation = TRUE)

y_dens <- matrix(0, N, d)
p_alpha <- matrix(0, N, d)
for (i in 1:N) {
  y_dens[i, ] <- dat$y[i, ] / sum(dat$y[i, ])
  p_alpha[i, ] <- exp(dat$mu + dat$zeta[i, ]) / sum(exp(dat$mu + dat$zeta[i, ]))
}

simPlotData <- data.frame(species=as.factor(rep(1:d, each=N)),
                          count=c(y_dens),
                          depth=rep(dat$X, times=d), alpha=c(p_alpha))

gsim1 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25)

gsim2 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth by species") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25) +
  facet_wrap( ~ species, ncol = 4)
multiplot(gsim1, gsim2, cols=1)
```



## dirichlet-multinomial simulation using Gaussian process without overdispersion

```{r}
dat <- sim_compositional_data(N=N, d=d, likelihood = "dirichlet-multinomial",
                              function_type = "gaussian-process",
                              correlation_function = "gaussian",
                              additive_correlation = FALSE,
                              multiplicative_correlation = TRUE)

y_dens <- matrix(0, N, d)
p_alpha <- matrix(0, N, d)
for (i in 1:N) {
  y_dens[i, ] <- dat$y[i, ] / sum(dat$y[i, ])
  p_alpha[i, ] <- exp(dat$mu + dat$zeta[i, ]) / sum(exp(dat$mu + dat$zeta[i, ]))
}

simPlotData <- data.frame(species=as.factor(rep(1:d, each=N)),
                          count=c(y_dens),
                          depth=rep(dat$X, times=d), alpha=c(p_alpha))

gsim1 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25)

gsim2 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth by species") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25) +
  facet_wrap( ~ species, ncol = 4)
multiplot(gsim1, gsim2, cols=1)
```



## dirichlet-multinomial simulation using Gaussian process without overdispersion

```{r}
dat <- sim_compositional_data(N=N, d=d, likelihood = "dirichlet-multinomial",
                              function_type = "gaussian-process",
                              correlation_function = "gaussian",
                              additive_correlation = FALSE,
                              multiplicative_correlation = TRUE)

y_dens <- matrix(0, N, d)
p_alpha <- matrix(0, N, d)
for (i in 1:N) {
  y_dens[i, ] <- dat$y[i, ] / sum(dat$y[i, ])
  p_alpha[i, ] <- exp(dat$mu + dat$zeta[i, ]) / sum(exp(dat$mu + dat$zeta[i, ]))
}

simPlotData <- data.frame(species=as.factor(rep(1:d, each=N)),
                          count=c(y_dens),
                          depth=rep(dat$X, times=d), alpha=c(p_alpha))

gsim1 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25)

gsim2 <- ggplot(simPlotData, aes(x=depth, y=count, color=species, group=species)) +
  geom_point(alpha=0.25) + theme(legend.position="none") +
  ggtitle("Simulated response vs. depth by species") +
  geom_line(aes(x=depth, y=alpha, col = species), simPlotData, lwd=1.25) +
  facet_wrap( ~ species, ncol = 4)
multiplot(gsim1, gsim2, cols=1)
```
